---
title: "Progress with the NOSAMS"
subtitle: "Hybrid Gas Ion Source"
author: "Brett Longworth"
institute: "NOSAMS/WHOI"
date: "2021/05/03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(xaringanthemer)

# Colors
dark_blue <- "#041e42"
light_blue <- "#00a9e0"
medium_blue <- "0069b1"
aqua_blue <- "#00b7bd"
cool_gray_11 <- "#53565a"
cool_gray_4 <- "#bbbcbc"
background <- "#e6e7e8"

style_mono_light(
  # colors
  base_color = dark_blue,
  # primary_color = background,
  # secondary_color = dark_blue,
  
  # fonts
  header_font_google = google_font("Montserrat", "800"),
  text_font_google = google_font("Red Hat Display", "500")
)
style_extra_css(
  css = list(
    ".title-slide" = list(
      "background-image" = "url(images/whoi_seal_title_bkgnd.jpg)",
      "background-size" = "cover"
    )
  ),
  append = TRUE,
  heading = "Extra CSS"
)
```

# Why run gas instead of graphite?

Benefits

- Skip a time-consuming reduction to graphite
- Eliminate blank associated with graphitization

Challenges

- Ultimate precision is lower
- Lower currents/counts
- variability in blank
- variability in sample performance

---

# How do you sputter a gas?

introduction onto Ti surface in target recess
Predicted chemistry/physics of decomposition and ionization

$$CO2+e^- \longrightarrow C^-+2O^-$$
* Inject CO<sup>2</sup> gas onto Ti target
* Sputter with Cs<sup>+</sup>
* Dissociate molecules (hopefully)
* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
* Extract C<sup>-</sup> beam

figure: Salazar target model

---

# Method history

Middleton?
NEC: Xu 2006
ETH:
source and inlet refinement

State of the art:
- reproducibilty: 1%
- size: >= 10ug C
- speed: XX minutes per sample or samples/day

---

# The NEC HGIS

figure: NEC arm drawing
photo: source 

* Modified to allow easy switch between solid and hybrid operation
* Rest is gas handling

Figure/photo of NEC bellows system

---

# Targets

.pull-left[
Potential for:

* Lower cost
* Better performance
]

.pull-right[
photo: target cutaway with callouts
]
---


# First tests - setup

Direct CO2 from bottle
Control flow by pressure and glass capillary

Conditions
* Modern and dead CO 2
* 8m x 50 μm capillary
* 0-4 atm supply pressure

Results
* ~10μA 12 C -
* <1% precision
* 1% blank
* 1-3% total efficiency

---

# First tests - results

.pull-left[
current vs flow
ratio vs flow
efficiency
precision

```{r}
names(mtcars)
```
]

.pull-right[
```{r, echo=FALSE}
ggplot(mtcars, aes(disp, mpg)) + geom_point() + theme_xaringan()
```
]

---

# Inlet systems

$$PV=nRT$$

AKA interfaces

Closed vs Open
Introduce steady, reproducible current as key to data quality

Dillution
Makes moving small amounts of gas easier by increasing pressure
Source doesn't care about helium as long as it gets its CO2
pumping issues

Gas delivered at set rate

For a given sample size:
* High inlet volume = low inlet pressure (NEC inlet)
* Low inlet volume = high inlet pressure (Ionplus inlet)


Smallest samples require minimizing inlet volume and pressure while maintaining good flow control

---

# Inlet systems - closed inlet

.pull-left[
IRMS dual bellows

* VG Prism dual bellows
* LabVIEW interface
* 2 x bellows
* Cryotraps
* Manifold capable
]

.pull-right[
figure: irms inlet
photo: irms inlet
]

---

# Closed inlet results

plot: Results from breakseals

---

# Open split development

development

photo: gas panel

---

# Current open split

current layout

photo: open split
figure: open split

---

# Carbonate method

Produce sample
target cleaning
introduce sample using helium

photo: gilson

---

# Results - current profiles

plot: long runs from vial

---

# Results - blanks

Current/target dependent?
Constant blanks and currents allow single mass independent correction

plot: current dependent blanks with intersect

---

# Results - standards

precision
reproducibility

plot: gas standards

reproducibility of straight gas vs vial standards


---

# Results - secondaries

precision, accuracy

plot: secondary standard performance

issues with blank and normalization

---

# Sample size reduction

high dillution, high flow
pressure control to maintain constant CO2 to source

plot: dillution modelling plot
figure: setup with pressure control

---

# MICADAS

source differences
* one target in source at a time
* gas supplied via conical joint in side of target

photo: micadas source

---

# MICADAS - Carbonate method

inlet/method differences
* micadas uses closed system
* co2 from carbonates transferred onto resin trap in stream of he
* sample introduced by adding helium and adjusting flow using variable volume

---

# Summary

---

# NEC/Ionplus inlets

compare high and low volume approach
