---
title: "Progress with the NOSAMS"
subtitle: "Hybrid Gas Ion Source"
author: "Brett Longworth"
institute: "NOSAMS/WHOI"
date: "2021/05/03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(xaringanthemer)
style_mono_light(base_color = "#23395b")
```

# Why run gas instead of graphite?

- Skip a time-consuming reduction to graphite
- Eliminate blank associated with graphitization

- Ultimate precision is lower
- Lower currents/counts
- variability in blank
- variability in sample performance

---

# How do you sputter a gas?

introduction onto Ti surface in target recess
Predicted chemistry/physics of decomposition and ionization

* Inject CO<super>2</super> gas onto Ti target
* Sputter with Cs<super>+</super>
* Dissociate molecules (hopefully)
* Produce C - ions (and O - , CH - , OH - , CO - ....)
* Extract C - beam

figure: Salazar target model

---

# Method history

Middleton?
NEC: Xu 2006
ETH:
source and inlet refinement

State of the art:
- reproducibilty: 1%
- size: >= 10ug C
- speed: XX minutes per sample or samples/day

---

# The NEC HGIS

NEC arm drawing
Target and source images
Figure/photo of NEC bellows system

* Modified to allow easy switch between solid and hybrid operation
* Rest is gas handling

---

# Targets

target cutaway with callouts

Potential for:

* Lower cost
* Better performance

---


# First tests - setup

Direct CO2 from bottle
Control flow by pressure and glass capillary

Conditions
* Modern and dead CO 2
* 8m x 50 μm capillary
* 0-4 atm supply pressure

Results
* ~10μA 12 C -
* <1% precision
* 1% blank
* 1-3% total efficiency

---

# First tests - results

current vs flow
ratio vs flow
efficiency
precision

```{r}
names(mtcars)
ggplot(mtcars, aes(disp, mpg)) + geom_point() + theme_xaringan()
```

---

# Inlet systems

$$PV=nRT$$

AKA interfaces

Closed vs Open
Introduce steady, reproducible current as key to data quality

Dillution
Makes moving small amounts of gas easier by increasing pressure
Source doesn't care about helium as long as it gets its CO2
pumping issues

Gas delivered at set rate

For a given sample size:
* High inlet volume = low inlet pressure (NEC inlet)
* Low inlet volume = high inlet pressure (Ionplus inlet)


Smallest samples require minimizing inlet volume and pressure while maintaining good flow control

---

# Inlet systems - closed inlet

IRMS dual bellows

Diagram
Photo

* VG Prism dual bellows
* LabVIEW interface
* 2 x bellows
* Cryotraps
* Manifold capable

---

# Closed inlet results

Results from breakseals figure

---

# Open split 

development
current layout
photo
diagram

---

# Carbonate method

Produce sample
target cleaning
introduce sample using helium

---

# Results - current profiles

figure: long runs from vial

---

# Results - blanks

Current/target dependent?
Constant blanks and currents allow single mass independent correction
figure: current dependent blanks with intersect

---

# Results - standards

precision
reproducibility

figure: gas standards

---

# Results - secondaries

precision, accuracy
figure: secondary standard performance

issues with blank and normalization

---

# Sample size reduction

high dillution, high flow
pressure control to maintain constant CO2 to source
figure: dillution modelling plot
figure: setup with pressure control

---

# MICADAS

source differences
* one target in source at a time
* gas supplied via conical joint in side of target

photo: micadas source

---

# MICADAS - Carbonate method

inlet/method differences
* micadas uses closed system
* co2 from carbonates transferred onto resin trap in stream of he
* sample introduced by adding helium and adjusting flow using variable volume

---

# Summary

---

# NEC/Ionplus inlets

compare high and low volume approach
