---
title: "HGIS flow calcs"
output:
  html_document: default
  html_notebook: default
---


```{r}
source('~/Projects/HybridGIS/R/HGISFunctions.R')
```

# Calculated Flow rates

To model laminar flow in a capillary, Pouiselle's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

  The functions loaded above calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{min}$ and pressure in kPa.

## Initial NOSAMS parameters

Initial tests used a 0.050mm x 8.2m deactivated silica caplillary. For this
capillary, parameters for Pouiselle's equation are as follows:

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2
flowcalc(300, r, u, x)
```

## Parameters for published hybrid GIS systems

Calculate flow for Xu et al 2006...

```{r}
flowcalc(1.500, 12.5E-5, u, 1.25)

```
Calculate flow for ETH...

```{r}
flowcalc(15.6, 3E-5, u, 2)
flowcalc(31.3, 3E-5, u, 2)

```
Calculate flow for CAMS...

```{r}
flowcalc(42, 7.5E-5, u, 6)

```
Calculate flow for Oxford...

```{r}
flowcalc(200, 3E-5, u, 1.25)

```

## Parameters for next USAMS tests

Short capillary at 1 atm

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 3
flowcalc(100, r, u, x)
```

Low pressure 15 mbar
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 1.0
flowcalc(1.5, r, u, x)
```
Medium pressure 150mbar
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 1.5
flowcalc(15, r, u, x)
```

Medium pressure 150mbar 100um x 1.58m PEEK capillary
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(1.5, r, u, x)
```

