---
title: "Open Split Tests"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Load libraries and functions

```{r}
library(tidyverse)
library(ggplot2)
library(amstools)
library(here)
source(here('/R/HGISFunctions.R'))
```

Development notes for using the HGIS with an open split. Intention is to use
with gasbench for carbonates and/or large co2 samples.

## Capillary parameters

Calculate ideal capillary to provide ~7.5uL/min at atmospheric pressure.

### Short capillary at 1 atm

40um capillary

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 3.5
flowcalc(100, r, u, x)
```
```{r}
r <- 2E-5
u <- 1.49E-5
x <- 2
flowcalc(100, r, u, x)
```
50um capillary
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 3
flowcalc(100, r, u, x)
```
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 7
flowcalc(100, r, u, x)
```

## First tests

### DATE?

Tried a bunch of capillary combinations. Other than initial pressure burst when
opening or closing HGIS valve, couldn't see any source pressure changes or beam
current. We believe the source end of the stainless capillary was clogged by
cesium residue. 

Cleaned capillary end at next source cleaning, first with kimwipe and water,
then with isopropanol. Lots of black crud. Confirmed open passage by blowing N2
through capillary into cup of isopropanol to observe bubbles.

### May 23, 2018

Tried 3m of 50um glass capillary. Insulated from source to cage with
teflon tubing as outer jacket. No problems with sparks.

Low currents, even when cutting capillary down to 2m. Either steel
capillary is partially clogged or tip is not making good seal with target.

### May 31, 2018 - He Dilution

Open split set up with helium dilution. Using 100um glass capillary. ~50cm
source to bulkhead with teflon jacket, 285cm bulkhead to open split. 

Table: Source pressures

Condition     Torr  
---------     ---- 
base w/cryo   1E-7  
base w/turbo  4E-7  
capillary     36E-7 


#### Dilution and current

Flow and current tests with varying capillary and dilution. Flows in mL/min,
pressure in Torr, current in uA. Replumbed helium to panel open split with
needle valve partway through to better control flow. Live CO2 set to 10mL/min.

```{r}
co2 <- c(18, 18, 18, 18, 18, 10, 10, 10, 10, 10, 10, 10, 10, 10, 10, 10)
caplen <- c(322, 322, 322, 322, 322, 322, 322, 322, 322, 322, 322, 322, 230, 230, 177, 177)
helium <- c(0, 100, 200, 300, 500, 0, 10, 20, 30, 50, 100, 200, 100, 200, 100, 50)
sourcevac <- c(7.5E-6,1.5E-6,1E-6,8.5E-7,8.4E-7,7.5E-7, 3.3E-6, 2.3E-6, 1.9E-6, 1.5E-6, 1E-6, 9.3E-7, 1.1E-6, 1.4E-6, 1.7E-6, 2.3E-6)
current <- c(7, 4, 3, 2, 2, 6.7, 5, 4, 3.5, 2.6, 1.6, 1.3, 1.3, 1.8, 2.3, 3.6)

osdil <- data.frame(co2, caplen, helium, sourcevac, current)
write.csv(osdil, here("Data/dillution.csv"))
osdil <- mutate(osdil, ratio = helium/co2)
osdil
```

```{r}
filter(osdil, co2 == 18) %>%
  ggplot(aes(ratio, current)) +
    geom_line()
filter(osdil, co2 == 10, caplen == 322) %>%
  ggplot(aes(ratio, current)) +
    geom_line()
```


#### Test Acquisitions

Ran 4 test Acquisitions with open split: pure CO2, pure dead CO2, dilluted
live, diluted dead.

```{r}
# Load data

data <- readResfile("/mnt/shared/USAMS/Results/2018 Results/USAMS053018R.txt")
data <- mungeResfile(data)
sdata <- data %>% 
  filter(Sample.Name != "Ceylon") %>%
  group_by(Sample.Name) %>%
  summarise(mean = mean(cor1412he),
            sd = sd(cor1412he),
            se = se(cor1412he),
            psd = sd(cor1412he)/mean,
            pse = se/mean,
            interr = 1/sqrt(sum(CntTotGT)),
            N = n())
sdata
```

Normalize!

```{r}
sdata %>% mutate(normfm = mean/9.8669966 * 1.0398)
```

```{r}
ospure <- filter(data, Sample.Name == "LiveGas")

ggplot(ospure, aes(ts, cor1412he)) + geom_point()
summary(ospure$cor1412he)

mean(ospure$cor1412he)
sd(ospure$cor1412he)
```

Load pos 7 (new target). Initial current 20uA (lots of LE13C - hydrides).
Down to 2uA in 2 min. 10mL/min of pure modern CO2 into open split.
Acquired 60 x 30s blocks.

PLOT

```{r}

```


### May 31, 2019 - startup tests

Reinstalled gas delivery arm after cleaning and polishing ball end.

Using 50um x .5m capillary source to cage and 50um x 2.5m to open split. Open split flowing modern co2 at 10mL/min. Using same targets and wheel as last time. Surprisingly good performance, including no-gas base current, hgis current and solid samples.


### Jun 3 2019 - dillution tests

Spun up inlet turbopump.