---
title: "Open Split Tests"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Load libraries and functions

```{r}
library(tidyverse)
source('~/Projects/HybridGIS/R/HGISFunctions.R')
```

Development notes for using the HGIS with an open split. Intention is to use
with gasbench for carbonates and/or large co2 samples.

## Capillary parameters

Calculate ideal capillary to provide ~7.5uL/min at atmospheric pressure.

### Short capillary at 1 atm

40um capillary

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 3.5
flowcalc(100, r, u, x)
```
```{r}
r <- 2E-5
u <- 1.49E-5
x <- 2
flowcalc(100, r, u, x)
```
50um capillary
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 3
flowcalc(100, r, u, x)
```
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 7
flowcalc(100, r, u, x)
```
## First tests

DATE?

Tried a bunch of capillary combinations. Other than initial pressure burst when
opening or closing HGIS valve, couldn't see any source pressure changes or beam
current. We believe the source end of the stainless capillary was clogged by
cesium residue. 

Cleaned capillary end at next source cleaning, first with kimwipe and water,
then with isopropanol. Lots of black crud. Confirmed open passage by blowing N2
through capillary into cup of isopropanol to observe bubbles.

